---
title: "Network analysis of cocaine data"
output: html_document
---

Load libraries.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(Hmisc) 
library(pheatmap)
library(RColorBrewer)
library(dendextend)
```

```{r saving functions and settings}
source('network_analysis_helper_functions.R')

saving_dir="files/"
save=FALSE #switch off for testing
```

Load data
```{r loading}
filtered_data<- read_csv("files/addictive_data.csv", col_types="cdff") %>%
  bind_rows(read_csv("files/natural_data.csv", col_types="cdff"))

categories<-read_csv("files/ALL_Lukasz-kategorie.csv", show_col_types = FALSE) %>%   
  mutate(category=factor(category, levels=c("CTXpl","OLF", "HPF", "CTXsp", "STR", "PAL","TH", "HY", "MB", "P", "MY"))) %>%
  arrange(category) %>%
  semi_join(filtered_data)

```

clean up, and calculate correlation matrices for each group separately.

First, extract group labels. Next, reformat data into a list of 6 dataframes corresponding to each group and calculate correlation matrices.
```{r make list of groups}
#names
group_labels<-unlist(group_keys(group_by(filtered_data, group_label)))

#list of dataframes for each group
lista.danych <- filtered_data %>%
  #left_join(categories, by="abbrev" ) %>%
  mutate(case_id=as.character(case_id)) %>%
  group_by(group_label, .drop=T) %>%
  group_split(.keep=F) %>%
  set_names(group_labels) %>%   
  lapply( function(x) column_to_rownames(pivot_wider(x, names_from = abbrev, values_from=signal), var="case_id"))
  
#correlations
lista.corr <- lapply(lista.danych, function(x) rcorr(as.matrix(x), type="pearson")$r) #spearman byłby lepszy?

#reorder correlations according to anatomical order
correlations.within.categories<-lapply(lista.corr, function(x) {
  kolejnosc<-categories$abbrev
  #kolejnosc<-kolejnosc[order(match(kolejnosc, lookup$abbrev))]
  x[kolejnosc, kolejnosc]
})

categories_for_labeling <- data.frame(category=categories$category)
row.names(categories_for_labeling)<-categories$abbrev
```

Create anatomical heatmaps (that is, arranged by the location of the brain region) for each experimental group separately.
```{r plot anatomical heatmaps, fig.width=8,fig.height=8}



#create heatmaps
generate_anatomical_heatmap<-function(grupa) {
  my_heatmap<-pheatmap(correlations.within.categories[[grupa]], 
                     cluster_cols = F, cluster_rows = F,
                     color= colorRampPalette(rev(brewer.pal(n = 7, name =  "RdBu")))(100),
                     border_color=NA,
                     legend = T,
                     breaks = seq(-1, 1, length.out = 101),
                     legend_breaks = c(-1,0,1),
                     annotation_row = categories_for_labeling, 
                     cellheight=3, cellwidth = 3,
                     show_colnames = T, 
                     show_rownames = T,
                     fontsize = 10,
                     fontsize_col = 2.5,
                     fontsize_row = 2.5,
                     main = grupa)
}

anatomical<-lapply(group_labels, generate_anatomical_heatmap)
names(anatomical)<-group_labels
  
if(save) {
  lapply(group_labels, function(x) {
    save_pheatmap_png(anatomical[[x]], paste(saving_dir, x, sep=""), "-korelacje-anatomiczne", width=5000, height=5500, res = 280)})
}
```


We want to look at how structures cluster together according to correlations, not anatomy. For this, complete euclidian distances between rows of matrix are calculated and then grouped using the hierarchical clustering algorithm.
```{r clustered heatmap, fig.width=8,fig.height=8}
distances<-lapply(lista.corr, function(x) {dist(x)})

generate_distances_heatmap<-function(grupa, distances) {
  my_heatmap<-pheatmap(as.matrix(distances[[grupa]]),
                       clustering_distance_rows=distances[[grupa]],
                       clustering_distance_cols=distances[[grupa]],
                       color = colorRampPalette(brewer.pal(n = 7, name =  "RdYlBu"))(100),
                       breaks = seq(0, 20, length.out = 101), #manual scale to have all groups look the same
                       border_color=NA,
                       cellheight=3,
                       cellwidth = 3,
                       show_rownames = T, fontsize_row = 2.5, 
                       show_colnames = T, fontsize_col = 2.5,
                       treeheight_row = 0,
                       treeheight_col = 90,
                       fontsize=8,
                       main = grupa)
  
}


mapy<-lapply(group_labels, generate_distances_heatmap, distances=distances)
names(mapy)<-group_labels
drzewa<-lapply(mapy, function(x) {as.dendrogram(x$tree_row)})

if(save) {
  lapply(group_labels, function(x) {save_pheatmap_png(mapy[[x]], paste(saving_dir, x, sep=""), "-korelacje-podobienstwo", width=5000, height=5800, res = 300)})
}
```

This data can be more conveniently presented as a dendrogram.
```{r generate list of heights vs clusters vectors}
hpk<-lapply(drzewa, function(x) {heights_per_k.dendrogram(x)})
```

One way to describe a dendrogram quantitatively is to plot the number of modules vs relative height. The differences between experimental groups become clearly visible.
```{r reshape, plot and save}
ad=c('Saline', 'Cocaine1d', 'Cocaine7d')
nat=c('Water', 'Sugar1d', 'Sugar7d')

#convert to data table
hpk_dt<-rbindlist(lapply(group_labels, function(x) {data.table(height=hpk[[x]], k=as.numeric(names(hpk[[x]])), group=x )}))
hpk_dt[,relative_height:=height/max(height), by=group]

#plot
ga<-ggplot(data=hpk_dt[relative_height>0.2 & group %in% ad], aes(x=relative_height, y=k, color=group))+ geom_line(size=1)
ga
gn<-ggplot(data=hpk_dt[relative_height>0.2 & group %in% nat], aes(x=relative_height, y=k, color=group))+ geom_line(size=1)
gn



hpk_wide<-dcast(hpk_dt, k ~ group, value.var = "height")
hpk_wide_odwrotnie<-dcast(hpk_dt, height ~ group, value.var = "k")
hpk_wide_relative<-dcast(hpk_dt, k ~ group, value.var = "relative_height")
hpk_wide_odwrotnie_relative<-dcast(hpk_dt, relative_height ~ group, value.var = "k")
#save
if(save) {
  ggsave(paste(saving_dir, "ad-module-vs-height.png", sep=""), ga, "png")
  ggsave(paste(saving_dir, "nat-module-vs-height.png", sep=""), gn, "png")
  write.csv(hpk_wide_odwrotnie_relative, paste(saving_dir, "heights_rel.csv", sep=""), row.names = F)
  }

```

Find the number of modules af half of the tree height and save to csv files.
```{r}
modules_half_height<-lapply(drzewa, function(x) {
  cutree_1h.dendrogram(x, 0.5*get_branches_heights(x, decreasing = T)[1], order_clusters_as_data = F)})
#save
lapply(group_labels, function(x) {
  write.csv(modules_half_height[[x]], paste(saving_dir,x, '_modules_half_height.csv', sep=''), row.names = T)})
```
Plot the dendrograms - dendextend package
```{r, fig.width=12,fig.height=6}
modules_number<-lapply(modules_half_height, function(x) {max(x)})

paleta_kolorow=brewer.pal(9, 'Dark2')
lapply(group_labels, function(x) {
    png(paste(saving_dir, x,"-tree.png", sep=""),  width = 1200, height = 600)

    p<-drzewa[[x]] %>%
    set("branches_k_color", value = paleta_kolorow, k=modules_number[[x]]) %>% 
    set("branches_lwd", 1.2) %>%
    set("labels_cex", 0.4) %>%
    set("labels_colors", value = paleta_kolorow, k=modules_number[[x]])%>%
    plot(main=x)
    p
    dev.off()
  })

```