---
title: "GLM with negative binomial"
output: html_document
---

Code for comparing brain imaging data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(MASS)
library(multcomp) #for glht - general linear hypothesis test
library(rstatix) #for multiple comparison correction

```

Data is loaded from csv files in long format, with abbrev being abbreviated name of the brain region as in Allen atlas.  

```{r}
categories<-read_csv("files/ALL_Lukasz-kategorie.csv", show_col_types = FALSE) %>%   mutate(category=as.factor(category))
addictive<- read_csv("files/addictive_data.csv", col_types="cdff") %>% left_join(categories )
natural<-  read_csv("files/natural_data.csv", col_types="cdff") %>% left_join(categories )
```

I fit a negative binomial model structure by structure. The Saline group is treated as reference.

```{r model group by group addictive data}
suppressWarnings(nb.addictive<-  addictive %>% 
  group_by(abbrev) %>%
  group_modify(~ broom::tidy(glm.nb(signal~group_label, data=.x))))
nb.addictive.wyniki<-nb.addictive %>% 
    mutate(term=gsub("group_label", "", term)) %>%
    filter(term != "(Intercept)")
```

For "natural" experiment, Water group is reference

```{r model group by group natural data}
suppressWarnings(nb.natural<-  natural %>% 
  group_by(abbrev) %>%
  group_modify(~ broom::tidy(glm.nb(signal~group_label, data=.x))))
nb.natural.wyniki<-nb.natural %>%
      mutate(term=gsub("group_label", "", term)) %>%
      filter(term != "(Intercept)")
```

##Testing the results 

The standard multiple comparison procedure to address the many-to-one problem (ie, comparing to a single control group) is the Dunnett  test. In our case, there are m=2 comparisons.

```{r addictive Dunnett test}
suppressWarnings(addictive.D.test<-  addictive %>% 
  group_by(abbrev) %>%
  group_modify(~ broom::tidy(multcomp::glht(glm.nb(signal~group_label, data=.x),
                     linfct = mcp(group_label = "Dunnett"),alternative = "greater")) #choice of two-sided causes p to double.
               ) %>%  mutate(null.value=NULL, term=NULL) 
)

addictive.with.D<-nb.addictive.wyniki %>%
  mutate(estimate=NULL, std.error=NULL, statistic=NULL) %>%
 left_join(addictive.D.test %>% 
             mutate(term=gsub(" - Saline", "", contrast), .keep="unused") %>% 
             mutate(estimate=NULL, std.error=NULL, statistic=NULL),
           by=c("abbrev", "term"))
```

```{r natural Dunnett test}
suppressWarnings(natural.D.test<-  natural %>% 
  group_by(abbrev) %>%
  group_modify(~ broom::tidy(multcomp::glht(glm.nb(signal~group_label, data=.x),
                     linfct = mcp(group_label = "Dunnett"),alternative = "greater")) #choice of two-sided causes p to double.
               ) %>%  mutate(null.value=NULL, term=NULL) 
)

natural.with.D<-nb.natural.wyniki %>%
  mutate(estimate=NULL, std.error=NULL, statistic=NULL) %>%
 left_join(natural.D.test %>% 
             mutate(term=gsub(" - Water", "", contrast), .keep="unused") %>% 
             mutate(estimate=NULL, std.error=NULL, statistic=NULL),
           by=c("abbrev", "term"))
```

Next, multiple comparison corrections are applied.

```{r FDR corrections}
addictive.with.fdr<-addictive.with.D %>% 
  left_join(categories ) %>% 
  relocate(category, .after=abbrev) %>%
  group_by(term) %>%
  rstatix::adjust_pvalue(p.col = "adj.p.value", output.col="p.fdr", method="fdr") %>%
  relocate(term, .after=category) 

natural.with.fdr<-natural.with.D %>% 
  left_join(categories )%>% 
  relocate(category, .after=abbrev) %>%
  group_by(term) %>%
  rstatix::adjust_pvalue(p.col = "adj.p.value", output.col="p.fdr", method="fdr") %>%
  relocate(term, .after=category) 
```
```{r save}
write.csv(addictive.with.fdr, "files/addictive_data_binomial.csv", row.names=F)
write.csv(natural.with.fdr,"files/natural_data_binomial.csv",  row.names=F)
```

Optionally, generate boxplots for structures where the differences after corrections are significant
```{r}
boxplots.a<-addictive %>%
  filter(abbrev %in% (addictive.with.fdr %>% filter(p.fdr<=0.05) %>% pull(abbrev)))  %>% 
  group_by(category) %>%
  group_map(
    ~ggplot(data = .x, aes(x=abbrev, y=signal)) + 
      geom_boxplot(aes(fill=group_label))+
      scale_fill_brewer(palette = "Dark2") +
      labs(title=.y$category))
print(boxplots.a)
```

```{r}
boxplots.n<-natural %>%
  filter(abbrev %in% (natural.with.fdr %>% filter(p.fdr<=0.05) %>% pull(abbrev)))  %>% 
  group_by(category) %>%
  group_map(~ggplot(data = .x, aes(x=abbrev, y=signal)) + geom_boxplot(aes(fill=group_label))+
              scale_fill_brewer(palette = "Accent")+
              labs(title=.y$category))
print(boxplots.n)
```